---
title: "Workshop 3 - R Markdown and Workshop 4 - Data Wrangling"
author: "Suzanne Nicholas"
date: '2022-03-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)


```

### Data wrangling


```{r}

survey<-read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv")


country<-survey %>%
  distinct(country)

survey%>%count(country, sort=TRUE)

survey_short<-survey%>%
  filter(
    country=="United States" | 
      country=="USA" | 
      country=="US"|
      country=="U.S."|
      country=="United States of America"|
      country=="Usa"
    )


survey_short%>%filter(
  annual_salary>50000 &
    gender=="Woman"
)


survey_short%>%
  select(how_old_are_you, industry, annual_salary, gender)%>%
  filter(gender=="Woman")%>%
  count(industry, sort = TRUE)


```



```{r}

#data loading

data<-read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv")

data%>%head()


```

Arranging observations

```{r}



```



Groupings and new variables

```{r}

survey_short%>%
  mutate(total_salary=annual_salary+other_monetary_comp)%>%
  select(industry, total_salary, annual_salary, other_monetary_comp)


survey_short%>%
  mutate(average_salary=mean(annual_salary, na.rm=TRUE))%>%
  select(industry, average_salary)

survey_short%>%
  group_by(industry, gender)%>%
  mutate(average_salary=mean(annual_salary, na.rm=TRUE))%>%
  select(industry, gender, average_salary)%>%
  arrange(desc(average_salary))%>%
  ungroup()

survey_short%>%
  group_by(gender, industry)%>%
  mutate(average_salary=mean(annual_salary, na.rm=TRUE))%>%
  select(industry, gender, average_salary)%>%
  arrange(desc(average_salary))%>%
  ungroup()




#add a new variable max_salary which is maximum salary per industry

survey_short%>%
  group_by(industry)%>%
  mutate(max_salary=max(annual_salary, na.rm=TRUE))%>%
  select(industry, max_salary)%>%
  arrange(desc(max_salary))%>%
  ungroup()





```

summarising data

```{r}

survey_short%>%
  group_by(industry)%>%
  summarise(average_salary=mean(annual_salary, na.rm=TRUE))

#group by industry and gender and use summarise for average_salary

survey_short%>%
  group_by(industry, gender)%>%
  summarise(average_salary=mean(annual_salary, na.rm=TRUE))


#group by gender and industry and use summarise for average_salary

survey_short%>%
  group_by(gender, industry)%>%
  summarise(average_salary=mean(annual_salary, na.rm=TRUE))

#group by industry and gender and use summary for average_salary and max_salary

survey_short%>%
  group_by(industry, gender)%>%
  summarise(
    average_salary=mean(annual_salary, na.rm=TRUE), 
    max_salary=max(annual_salary, na.rm=TRUE)
    )
    

```


Case_when

```{r}

survey_short_veryshort<-survey_short%>%
  select(industry, gender, annual_salary)


#if annual_salary is greater than $700,000 than your annual_salary_level is high, otherwise it is low
# if annual_salary is more than $70,000, but less or equal to 700000 then your annual_salary_level is middle, otherwise it is low.

survey_short_veryshort%>%
  mutate(
    annual_salary_level=case_when(
      annual_salary>700000 ~"high",
      annual_salary>70000 & annual_salary<=700000 ~ "middle",
      T~"low"
    )
  )%>%
  arrange(desc(annual_salary))


#create a variable highly_paid_women
#if gender == "Woman" and annual_salary>50000 then (~) it is "high", otherwise it is low

survey_short_veryshort%>%
  mutate(
    highly_paid_women=case_when(
      gender=="Woman" & annual_salary>50000 ~ "high",
      T ~ "low"
    )
  )%>%
  
  filter(gender=="Woman")%>%
  arrange(desc(annual_salary))




```

```{r}

data(mpg)

mpg%>% ggplot(
  aes(displ, hwy, colour = class)) +
  geom_point()
  
mpg%>% ggplot(
  aes(displ, hwy, colour = year, size = cyl)) +
  geom_point()
  


```

#install.packages("skimr")

#library(skimr)

#skim(survey)

#task generate as many as possible BEAUTIFUL and MEANINGFUL data viz using survey dataset.

```{r}

#install.packages("skimr")

library(skimr)

skim(survey)


```



```{r}

computing_industry%>%ggplot(
  aes(how_old_are_you)) +
  geom_bar() %>%
  
  labs(
    title = "AGE OF TECH EMPLOYEES") +
  
  xlab("age") +
  ylab("no. of employees")

```

```{r}

computing_industry%>%ggplot(
  aes(gender) +
  geom_bar() +
  
  labs(
    title = "Mix of Genders in tech") +
  
  xlab("gender") +
  ylab("no. of employees")

```


```{r}

computing_industry%>%
  mutate(
    computing_salaries=case_when(
      (annual_salary+other_monetary_comp)>=200000 ~ "Greater than $200,000",
      (annual_salary+other_monetary_comp)>100000 & (annual_salary+other_monetary_comp)<200000 ~ "Between $100,000 and $199,999",
      T ~ "lower than $99,999"
    )
  )%>%
  
 ggplot(
   aes(computing_salaries, colour=gender)) +
  geom_bar(
 )+
  
  labs(
    title = "Salaries in tech") +
  
  xlab("salary ranges") +
  ylab("no. of employees")
 

```


usa%>%
   mutate(
    summary_salaries=case_when(
      (annual_salary+other_monetary_comp)>=200000 ~ "Greater than $200,000",
      (annual_salary+other_monetary_comp)>100000 & (annual_salary+other_monetary_comp)<200000 ~ "Between $100,000 and $199,999",
      T ~ "lower than $99,999"
    )
  )%>%
  
 ggplot(
   aes(summary_salaries, colour=how_old_are_you)) +
  geom_bar(
 )
 
```{r}

library(scales)
library(lubridate)

```
 
 

```{r}

survey <- survey %>%
  mutate(timestamp = mdy_hms(timestamp),
         age_category = fct_relevel(fct_reorder(how_old_are_you, parse_number(how_old_are_you)), "under 18"),
         experience_overall = str_replace(overall_years_of_professional_experience, " - ", "-"),
         experience_overall = fct_reorder(experience_overall, parse_number(experience_overall)),
         experience_in_field = str_replace(years_of_experience_in_field, " - ", "-"),
         experience_in_field = fct_reorder(experience_in_field, parse_number(experience_in_field))) %>%
  mutate(gender = fct_collapse(coalesce(gender, "Other or prefer not to answer"), "Other or prefer not to answer" = c("Other or prefer not to answer", "Prefer not to answer")),
         race = fct_lump(coalesce(race, "Other"), 4),
         age=parse_number(how_old_are_you))



```


```{r}

survey_short %>%
  ggplot(aes(annual_salary)) +
  geom_histogram() +
  scale_x_log10(labels = dollar_format()) +
  labs(x = "Annual")



```





